%both results
\chapter{Conclusion and Outlook} \label{sec:outlook}

%hier mehr aufs tempus achten!
This thesis presented two types of analyses in the field of neutrino point source searches of time-integrated and time-dependent nature using the software \texttt{csky} and gave insight into the difficulties of time-dependent stacking.
Both main analyses provided preliminary results on the contribution of neutrinos at neutrino alert positions of the gamma-ray-followup gold type.
$\num{9}$ years of IceCube data were used for this.
The simulation data has been cleaned of GFU-gold like events to prevent a bias, since all events of the GFU-gold like events counted as source events.

The first analysis is time-integrated and examined the stacked contribution of $\num{72}$ sources.
For this purpose, the implementation of the analysis was checked in the form of background and signal trials.
In addition, the fit behaviour of the spectral index of the signal emission and the number of signal events was examined.
Finally, the time-integrated analysis yielded a sensitivity of $\SI{1.68e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}$ and a $\num{5}\sigma$ discovery potential of $\SI{6.81e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}$, which are about one order of magnitude below the all-sky search \cite{all_sky_paper}.

The time-dependent analysis examined the contribution at $\num{10}$ selected source positions individually and is based on a new, conservative approach in which the temporal centre of the neutrino emission is fixed to the arrival time of the source event and the width can vary freely over a period of up to $\num{200}$ days.
Accordingly, fewer assumptions were made about the emission behaviour of the source, making the analysis even more conservative.
The resulting distributions of the time windows in signal and background trials were examined and additionally broken down in connection with the fitted number of signal events.
Here, again, the examination of the fit parameters provided indications of the correct implementation and functionality of the analysis.
The emerging preliminary results of sensitivity and discovery potential were presented in dependence of their spatial and temporal position.
The best results here were a sensitivity of $\SI{4.63e-2}{\giga\electronvolt\centi\meter\tothe{-2}}$ and a $\num{5}\sigma$ discovery potential of $\SI{1.32e-1}{\giga\electronvolt\centi\meter\tothe{-2}}$, which are in the range of a comparable analysis \cite{martina} and far below the total measurable flux of IceCube.

Finally, the difficulties of time-dependent stacking were discussed.
For this purpose, a model was developed for calculation, implemented on the \texttt{tdepps} software and checked for correct functionality.
However, the unusual results led to the use of \texttt{csky} for the computation of the main analyses in this thesis.

Improving the methods of point source searches is an important step towards the discovery of neutrino point sources and, accordingly, an important step towards the understanding of the processes inside cosmic objects.

%The time-integrated analysis provides a sensitivity of
%\begin{align}
%  \SI{1.68e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}
%\end{align}
%and a $\num{5}\sigma$ discovery potential of
%\begin{align}
%  \SI{6.81e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}
%\end{align}
%per source at a spectral index of $\gamma=\num{2}$ and a reference energy of $E_0=\SI{100}{\tera\electronvolt}$.
%This result is about one order of magnitude below the results of the all-sky scan \cite{all_sky_paper} referenced in figure \ref{fig:all_sky_plot}.
%The time-dependent analysis further provides a best sensitivity of
%\begin{align}
%  \SI{4.63e-2}{\giga\electronvolt\centi\meter\tothe{-2}}
%\end{align}
%and a $\num{5}\sigma$ discovery potential of
%\begin{align}
%  \SI{1.32e-1}{\giga\electronvolt\centi\meter\tothe{-2}}
%\end{align}
%for the GFU-gold event with id $\num{55820225}$ corresponding to source number $\num{58}$ also with a spectral index of $\gamma=2$ and at a reference %energy of $E_0=\SI{100}{\tera\electronvolt}$.
%A comparable analysis \cite{martina} provides a preliminary discovery potential at $\num{3}\sigma$ of
%\begin{align}
%  \SI{2.7e-2}{\giga\electronvolt\centi\meter\tothe{-2}}
%\end{align}
%The results in this thesis are slightly worse but are also in this region, considering that $\num{5}\sigma$ results are available here.
%
%It is also possible to compare these results with the diffuse astrophysical muon neutrino flux as is done, for example, in \cite{thorben}.
%Using the flux model from $\num{10}$ years of IceCube data \cite{flux_2},
%\begin{align}
%  \Phi^{\nu_\mu + \bar{\nu}_\mu} &= 1.44\cdot\left(\frac{E_\nu}{\SI{100}{\tera\electronvolt}}\right)^{-2.28}\cdot\SI{e-18}{\giga\electronvolt\tothe{-%1}\centi\meter\tothe{-1}\second\tothe{-1}\steradian\tothe{-1}},
%\end{align}
%a fluence upper limit can be generated.
%It is assumed that the entire flux comes from these $\num{10}$ sources in the form of a burst emission.
%The burst rate $R$ can be calculated with the total lifetime $T$ taken from table \ref{tab:data}
%\begin{align}
%  R &= \frac{N_\text{srcs}}{4\pi\cdot T}
%\end{align}
%and the fluence at a refernece energy of $\SI{100}{\tera\electronvolt}$ is about
%\begin{align}
%  E^2_\nu\Phi_0^{\nu_\mu + \bar{\nu}_\mu} &\approx \SI{1.44e-8}{\giga\electronvolt\centi\meter\tothe{-2}\second\tothe{-1}\steradian\tothe{-1}}.
%\end{align}
%The per-burst fluence with the burst time window $\Delta T$ canceling out then becomes
%\begin{align}
%  E^2_\nu\Phi_\text{GFU}^\text{diff} &= \frac{\Delta T \cdot E^2_\nu\Phi_0^{\nu_\mu + \bar{\nu}_\mu} \cdot 4\pi}{\Delta T \cdot R \cdot 4\pi},\\
%  &\approx \SI{4.98}{\giga\electronvolt\centi\meter\tothe{-2}}.
%\end{align}
%
%This limit is far above even the most insensitive result of $\SI{1.99e-1}{\giga\electronvolt\centi\meter\tothe{-2}}$, although the flux of $\num{10}$ %years is used for comparison and only $\num{9}$ years of data have been studied in this thesis.
%The result indicates that only a small fraction of the total flux originates from these $10$ source positions, and a larger result would mean that more %signal would be needed to identify these sources than has been measured so far.

\section{Discussion and Outlook}

The main goal of this thesis was to perform a point source search at neutrino alert positions, to investigate the difficulties of time-dependent stacking and to implement an experimental approach to overcome these problems.

The next step would be to unblind the results.
Although this step does not require any further work in the analysis, the unblinding process requires a group of reviewers from the IceCube collaboration to work meticulously through the analysis and question the analysis in a very time-consuming process.
At this stage, however, this step was beyond the scope of the Master's thesis and was therefore not be carried out.

In addition, the trial correction via post-trials is missing for the time-dependent search.
The reason for the trial correction is that only one source position was examined at a time on the basis of a GFU-gold event from a selected number of events, although the signal can also come from another GFU-gold event position.
The results are therefore actually somewhat worse than they initially appear.
An attempt was made to use the post-trial method presented in \cite{thorben}.
The strange course of the trial correction, which can be seen in the appendix in figure \ref{fig:post_trials}, indicates an incorrect implementation in this analysis.

It is also striking that in the simulation sample only about half the amount of GFU-gold events was identified and filtered out, which can be observed in figure \ref{fig:gfu_gold_comp}.
This could be due to the fact that antiparticles were not taken into account in the simulation files.
In addition, the events were identified using simulation sets 21002 and 21220.
However, these sets are impacted by a known bug.
Although the effects of this bug are expected to be small, it is advised to be cautious when comparing these sets to data.
This bug is addressed in the SnowStorm project \cite{snowstorm}.

The most striking observation of this work, however, is the strong scattering of the results of sensitivities and discovery potentials, which can be seen in figure \ref{fig:sens_disc_time_dep_dec}.
Normally, the sources would roughly follow a line corresponding to the all-sky survey, see figure \ref{fig:all_sky_plot}.
This scatter can be due to several causes.
One reason could be the spatial inaccuracy of the source positions, but this does not show a clear trend.
In addition, the analysis time windows could be in different data sets and even overlap two.
But even with this, no clear correlation can be discerned.
In addition, it is difficult to discern a certain pattern with only $\num{10}$ sources.
On the other hand, this analysis offers great insight into the behaviour of many analysis parameters.
These could help to explain the high result of source number $\num{37}$.
Many signal trials of this source failed, which can be observed in table \ref{tab:trials_sig_time_dep_table}.
This could explain the overall lower signal values in figure \ref{fig:sens_ns_dt_all}, which means that more signal is needed overall to achieve the desired threshold of sensitivity and discovery potential.
In general, more trials may ensure a better result.
Another argument could also be an outlier event that is scrambled unfortunately.
However, identifying such events is beyond the scope of this thesis.

% This is the start of the "outlook"
There are some aspects that can be improved in the analysis.
For instance, a more recent, improved point source dataset version with additional years of data can be used.
It would certainly also be interesting to set the spectral index of the source emission to that of the GFU-gold events.
Also the spatial uncertainties of the source positions could be accounted for.
Another point would be to investigate more than $\num{10}$ sources, which is certainly also possible later as stacking in \texttt{csky}.
A time-dependent stacking analysis would be more sensitive and would not require as much computing capacity as if all sources were examined individually.

In conclusion, this thesis provides a good basis for getting started with point source searches using \texttt{csky} and can be used for further analyses by making minor changes.
For example, the data set can be enlarged or additional or different source locations can be investigated.

%are the results measurable?\\
%time-int sens: $\SI{1.68e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}$ per source\\
%time-int disc $\num{5}\sigma$: $\SI{6.81e-14}{\tera\electronvolt\tothe{-1}\centi\meter\tothe{-1}\second\tothe{-1}}$ per source\\
%one magnitude under all-sky\\
%time-int total flux of $\SI{4.11}{\giga\electronvolt\centi\meter\tothe{-2}\second\tothe{-1}}$ calculated as thorben at $\SI{100}{\tera\electronvolt}$ %reference\\
%time-dep sens, best: $\SI{4.63e-2}{\giga\electronvolt\centi\meter\tothe{-2}}$ at $\SI{100}{\tera\electronvolt}$ reference\\
%time-dep disc $\num{5}\sigma$, best: $\SI{1.32e-1}{\giga\electronvolt\centi\meter\tothe{-2}}$ at $\SI{100}{\tera\electronvolt}$ reference\\
%time-dep weighted per burst fluence of $\SI{34.91}{\giga\electronvolt\centi\meter\tothe{-2}}$ calculated as thorben at $\SI{100}{\tera\electronvolt}$ %reference\\
%martina karl: disc $\SI{2.7e-2}{\giga\electronvolt\centi\meter\tothe{-2}}$ with $\num{3}\sigma$ but \textbf{what reference energy?}\\
%compare both analyses, \textbf{but how?}\\
%compare with other analyses\\

%what could have been done better, whats weird:\\
%- mc gfu gold events about times 2: 21002 21220 are weird\\
%- source 37 has few trials\\
%- more trials; timewindow distribution could be smoother\\
%- maybe theres a background event which is an outlier that gets badly scrambled\\
%- post trials should be done or all sources\\
%- different gammas\\
%- take gamma of gfu gold events\\
%- spatial errors of sources could be accounted for\\
%- take newest version of ps\_tracks\\
%why no unblinding?\\
%- takes too long\\
